---
title: "Descriptive analysis of binary variables in Lassa virus dataset"
---

Lassa "mock" data
```{r}
library(dplyr)
library(stringr)
library(purrr)
library(questionr)
library(naniar)
library(ggplot2)
library(Rmisc)
library(ggpubr)
library(tidyr)
library(textclean)
```

```{r}
filename = "C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Lassa_virus/practice_data/commcare-randomized-data-2018-2020.csv"

lassa <- read.csv(filename, header = TRUE, sep = ",")
```


Subset data to contain only clinical features before admission

```{r}
#Subset data
clinical_features.before_admission <- lassa %>%
  select(contains("clinical_features.before_admission")) %>%
  select(contains("duration")) %>%
  select(!contains("days")) 

#Remove redundant text from names
names(clinical_features.before_admission) <-  mgsub(names(clinical_features.before_admission), c(
  "clinical_features.before_admission_group.before_admission_symptom_duration.",
  "clinical_features.before_admission_group.before_admission_bleeding_question_group.bleeding_duration.",
  "_before_admission",
  "_duration.symptom"), c("", "", "", ""))
```

Now convert all values containing text to 1 and all empty values to 0 so we can treat these variables as binary.

```{r}
clinical_features.before_admission[clinical_features.before_admission != ""] <- 1
clinical_features.before_admission[clinical_features.before_admission == ""] <- 0
```

Combine some variables as described in Okokhere et al.
Severe CNS features = coma, seizure, tremors, confusion 
Non-severe CNS features = dizziness, lethargy, drowsiness (no lethargy or drowsiness in this dataset)

bleeding_all = eyes_bleeding, hematemesis (vomiting blood), hematochezia (blood in stool), hemoptysis (coughing up blood), injection_sites_bleeding, mouth_bleeding, nostril_bleeding, vaginal_bleeding, other_bleeding

```{r}
#subset bleeding variables and sum values in new variable, don't include eyes_bleeding
bleeding_names <- names(clinical_features.before_admission[c(2,4:9)])
bleeding <- clinical_features.before_admission[,bleeding_names]
bleeding <- map_df(bleeding, as.numeric)
bleeding$bleeding = rowSums(bleeding)
bleeding[bleeding > 1] <- 1

#subset severe cns variables and sum values in new variable
severe_cns_names <- c("coma", "seizure", "tremors", "confusion")
severe_cns <- clinical_features.before_admission[,severe_cns_names]
severe_cns <- map_df(severe_cns, as.numeric)
severe_cns$severe_cns = rowSums(severe_cns)
severe_cns[severe_cns > 1] <- 1
```

Make separate data frames for survived, died, and all patients with known outcome. 

```{r}
#vector of variables to include in final analysis - remove "red_eyes" because there are not
names <- c("hematuria", "abdominal_pain", "chest_pain", "cough", "deafness", "diarrhea", "dizziness", "dyspnoea", "fever", "headache", "jaundice", "myalgia", "neck_stiffness","restlessness", "sore_throat", "vomiting", "weakness", "other")

outcome <- lassa$clinical_outcome.outcome

cfba <- cbind("outcome" = outcome, "bleeding" = bleeding$bleeding, "severe_cns" = severe_cns$severe_cns, clinical_features.before_admission[,c(names)])
```

Calculate totals and percents of people who experienced each symptom for all, died, and survived.

```{r}
survived_bin <- cfba[grep("recovered", cfba$outcome), ]
died_bin <- cfba[grep("died", cfba$outcome), ]
all_bin <- rbind(survived_bin, died_bin)

#First save all the tables as variables - this gives the total number of patients experiencing each symptom
#select the second element of the table for each variable
all_total <- map(map(all_bin, table), 2)
S_total <- map(map(survived_bin, table), 2)
D_total <- map(map(died_bin, table), 2)

#Remove first row
all_total <- all_total[-1]
S_total <- S_total[-1]
D_total <- D_total[-1]
cfba_binary <- names(cfba)
cfba_binary <- cfba_binary[-1]

#Replace NULL values with o
all_total[all_total == "NULL"] <- 0
S_total[S_total == "NULL"] <- 0
D_total[D_total == "NULL"] <- 0

#Save percent data in variable

all_percent <- round(unlist(all_total)/nrow(all_bin), 2)*100
S_percent <- round(unlist(S_total)/nrow(survived_bin), 2)*100
D_percent <- round(unlist(D_total)/nrow(died_bin), 2)*100

#OI_percent_list <- map(map(all, function(x) prop.table(table(x))), 2)
#S_percent_list <- map(map(survived, function(x) prop.table(table(x))), 2)
#D_percent_list <- map(map(died, function(x) prop.table(table(x))), 2)

#unlist percent data, multiply by 100, round
#OI_percent <- round(unlist(OI_percent_list) * 100, 0)
#S_percent <- round(unlist(S_percent_list) * 100, 0)
#D_percent <- round(unlist(D_percent_list) * 100, 0)
```

Make strings to report data.
```{r}
all_string <- paste(all_total,"/",nrow(all_bin)," (",all_percent,"%)", sep="")
surv_string <- paste(S_total,"/",nrow(survived_bin)," (",S_percent,"%)", sep="")
died_string <- paste(D_total,"/",nrow(died_bin)," (",D_percent,"%)", sep="")
```

Reformat outcome variable as an ordered factor 

```{r}
all_bin[all_bin == "died"] <- 1
all_bin[all_bin == "recovered"] <- 0
all_bin <- data.frame(map(all_bin, as.factor)) #convert variables to factor
all_bin$outcome <- ordered(all_bin$outcome)
```

Create separate logistic regression model for each variable and calculate odds ratios by exponentiating coefficients.
```{r}
glm <- map(all_bin, function(x) glm(outcome ~ x, data = all_bin, family = binomial))
odds_df <- data.frame(map(glm[-1], coefficients))
odds_vector <- exp(as.vector(odds_df[-1,]))
odds_rounded_list <- as.list(round(odds_vector,2))

#Calculate p-values for odds ratios
p_value_list <- map(map(glm[-1], odds.ratio), "p")
p_value <- map(map(p_value_list, 2), function(x) round(x, 5))
```

Create a summary table of the data.

```{r}
binary_table <- data.frame(cbind(cfba_binary, all_string, surv_string,  died_string, as.character(p_value), as.character(odds_rounded_list)))
names(binary_table) <- c("Symptom", "Overall", "Survived", "Died", "p_value", "odds_ratio") 
binary_table <- binary_table %>%
  arrange(p_value)
binary_table
```

```{r}
p <- round(data.frame(unlist(p_value)), 6)
newdata <- cbind(cfba_binary, OP =  all_percent, SP =  S_percent, DP =  D_percent, p = p)
names(newdata) <- c("name", "Overall", "Survived", "Died",  "p")
```

Order according to increasing p value and generate bar plot
```{r}
newdata$name <- factor(newdata$name, levels = newdata$name[order(newdata$p)])

nd <- newdata %>%
  gather("Type", "Value", -name)

Type = c("Overall", "Died", "Survived", "p")
nd$Type <- factor(nd$Type, ordered = TRUE,  levels = Type)
```

```{r}
ggplot(subset(nd, Type %in% c("Overall" , "Died", "Survived")), aes(name, Value, fill = Type)) +
  geom_bar(position = "dodge", stat = "identity", colour="black") +
  scale_fill_manual(values=c("palegreen3", "coral2", "steelblue2")) +
  theme_bw() +
  ylab("Prevalence (%)") +
  xlab("Complication during treatment in hospital") +
  ylim(0, 100) +
  ggtitle("Prevalence of complications in patients diagnosed with Lassa fever")
```