---
title: "Validating LF prognostic model on a new dataset"
---

```{r}
library(dplyr)
library(textclean)
library(purrr)
library(ROCR)
library(caret)
```

I used this tutorial to guide my interpretation of logistic regression coefficients:
https://stats.idre.ucla.edu/other/mult-pkg/faq/general/faq-how-do-i-interpret-odds-ratios-in-logistic-regression/

Load data.

```{r}
filename = "C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Lassa_virus/practice_data/commcare-randomized-data-2018-2020.csv"

lassa_raw <- read.csv(filename, header = TRUE, sep = ",")
```

Only include observations where outcome is known.

```{r}
lassa <- lassa_raw %>%
  filter(clinical_outcome.outcome == "recovered" | clinical_outcome.outcome == "died")
```

Format binary variables.

```{r}
#Subset data
clinical_features.before_admission <- lassa %>%
  select(contains("clinical_features.before_admission")) %>%
  select(contains("duration")) %>%
  select(!contains("days")) 

#Remove redundant text from names
names(clinical_features.before_admission) <-  mgsub(names(clinical_features.before_admission), c(
  "clinical_features.before_admission_group.before_admission_symptom_duration.",
  "clinical_features.before_admission_group.before_admission_bleeding_question_group.bleeding_duration.",
  "_before_admission",
  "_duration.symptom"), c("", "", "", ""))

#convert all values containing text to 1 and all empty values to 0 so we can treat these variables as binary.
clinical_features.before_admission[clinical_features.before_admission != ""] <- 1
clinical_features.before_admission[clinical_features.before_admission == ""] <- 0
```

Combine some variables as described in Okokhere et al.
Severe CNS features = coma, seizure, tremors, confusion 

bleeding_all = eyes_bleeding, hematemesis (vomiting blood), hematochezia (blood in stool), hemoptysis (coughing up blood), injection_sites_bleeding, mouth_bleeding, nostril_bleeding, vaginal_bleeding, other_bleeding

```{r}
#subset bleeding variables and sum values in new variable, don't include eyes_bleeding
bleeding_names <- names(clinical_features.before_admission[c(2,4:9)])
bleeding <- clinical_features.before_admission[,bleeding_names]
bleeding <- map_df(bleeding, as.numeric)
bleeding$bleeding = rowSums(bleeding)
bleeding[bleeding > 1] <- 1

#subset severe cns variables and sum values in new variable
severe_cns_names <- c("coma", "seizure", "tremors", "confusion")
severe_cns <- clinical_features.before_admission[,severe_cns_names]
severe_cns <- map_df(severe_cns, as.numeric)
severe_cns$severe_cns = rowSums(severe_cns)
severe_cns[severe_cns > 1] <- 1
```

Subset variables from new dataset that were included in the original model.

```{r}
age <- lassa$basic_information.age
severe_cns <- severe_cns$severe_cns
bleeding <- bleeding$bleeding
jaundice <- clinical_features.before_admission$jaundice
ast <- lassa$blood_chemistry_list.aspartate_aminotransferase_ast.0
cr <- lassa$blood_chemistry_list.creatinine.0
k <- lassa$blood_chemistry_list.potassium.0

newdata <- data.frame(cbind(age, severe_cns, bleeding, jaundice, ast, cr,  k))
newdata <- map_df(newdata, as.numeric)
```

Validate prognostic model published in Okokhere et al. The Lancet (2018). The logistic regression coefficients are as follows:

Intercept = -8.540674959
Age = 0.043498988
Severe CNS = 1.011961038
Bleeding = 0.898388975
Jaundice = 2.029453703
AST = 0.002990266
Cr = 0.146072204
K = 0.923083855

Multiply each column by the corresponding coefficient and sum the values in each row in the final column. Ignore values of NA.

```{r}
age_val <- unlist(map(newdata$age, function(x) x * 0.043498988))
severe_cns_val <- unlist(map(newdata$severe_cns,  function(x) x * 1.011961038))
bleeding_val <-  unlist(map(newdata$bleeding, function(x) x * 0.898388975))
jaundice_val <- unlist(map(newdata$jaundice, function(x) x * 2.029453703))
ast_val <- unlist(map(newdata$ast,  function(x) x * 0.002990266))
cr_val <- unlist(map(newdata$cr, function(x) x * 0.146072204))
k_val <-  unlist(map(newdata$k, function(x) x * 0.923083855))

coef_df <- cbind.data.frame(age_val, severe_cns_val, bleeding_val, jaundice_val, ast_val, cr_val, k_val)

coef_sum <- rowSums(coef_df,  na.rm = TRUE)
coef_sum <- coef_sum + -8.540674959
```

Use the sum of the predictor values * coefficients to calculate the probability that outcome = 1 (died). If p > 0.5,  change value to died.

```{r}
num <- exp(coef_sum)
denom <- 1 + exp(coef_sum)
p <- num/denom
```

Graph ROC and calculate AUC.

```{r}
# format actual outcome values so died = TRUE
outcome <- ifelse(lassa$clinical_outcome.outcome == "died", TRUE, FALSE)
outcome <- factor(outcome, levels = c(TRUE,FALSE))

# format predicted outcomes so p > 0.5 = TRUE
outcome.pred <- ifelse(p > 0.5, TRUE, FALSE)
outcome.pred <- factor(outcome.pred, levels = c(TRUE, FALSE))

# generate a confusion matrix to evaluate accuracy of the model 
confusionMatrix(outcome.pred, outcome)
```


```{r}
roc_pred <- prediction(predictions = p  , labels = outcome)
roc_perf <- performance(roc_pred , "tpr" , "fpr")
plot(roc_perf,
     colorize = TRUE)

pdf("./lassa_fever_analysis/ROC_model_validation_symptoms_before_admission.pdf")
plot(roc_perf,
     colorize = TRUE)
dev.off()
```


```{r}
auc_ROCR <- performance(roc_pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
paste0("The AUC is ",round(auc_ROCR,2))
```


