---
title: "Validating LF prognostic models - model #1 (lab-full)"
---

```{r}
library(dplyr) 
library(textclean)
library(purrr)
library(ROCR)
library(caret)
library(tidyr)
```

Load data.

```{r}
filename = "C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Lassa_virus/practice_data/commcare-randomized-data-2018-2020.csv"

lassa_raw <- read.csv(filename, header = TRUE, sep = ",")
```

Only include observations where outcome is known.

```{r}
lassa <- lassa_raw %>%
  filter(clinical_outcome.outcome == "recovered" | clinical_outcome.outcome == "died")
```

Define variables in new data
```{r}
outcome <- lassa$clinical_outcome.outcome
age <- lassa$basic_information.age
ast <- lassa$blood_chemistry_list.aspartate_aminotransferase_ast.0
cre <- lassa$blood_chemistry_list.creatinine.0
k <- lassa$blood_chemistry_list.potassium.0
tbil <- lassa$blood_chemistry_list.total_bilirubin.0
plt <- lassa$hematology_list.platelet.0
wbc <- lassa$hematology_list.white_cell_count.0
bun <- lassa$blood_chemistry_list.blood_urea_nitrogen.0
hct <- lassa$hematology_list.haematocrit_concentration.0
lym <- lassa$hematology_list.lymphocytes.0
tpro <- lassa$blood_chemistry_list.total_protein.0

newdata_all <- cbind.data.frame(outcome, age, ast, cre, k, tbil, plt, wbc, bun, hct, lym, tpro)
```

###### Test model #1 (lab-full)

```{r}
# format data for model 1
newdata <- drop_na(newdata_all, c(age, ast, cre, k, tbil, plt, tpro))
intercept <- -5.736604643 
age <- 0.035540840 
ast <- .003186223 
cre <- 0.130760630 
k <- 0.914579133 
tbil <- 0.015229333 
plt <- 0.001630707
tpro <- -0.353872521

age_val <- unlist(map(newdata$age, function(x) x * age))
ast_val <- unlist(map(newdata$ast, function(x) x * ast))
cre_val <- unlist(map(newdata$cre, function(x) x * cre))
k_val <- unlist(map(newdata$k, function(x) x * k))
tbil_val <- unlist(map(newdata$tbil, function(x) x * tbil))
plt_val <- unlist(map(newdata$plt, function(x) x * plt))
tpro_val <- unlist(map(newdata$tpro, function(x) x * tpro))

coef <- cbind.data.frame(age_val, ast_val, cre_val, k_val, tbil_val, plt_val, tpro_val)

coef_sum <- rowSums(coef, na.rm = TRUE)
coef_sum <- coef_sum + intercept
```

```{r}
num <- exp(coef_sum)
denom <- 1 + exp(coef_sum)
p <- num/denom
```

Graph ROC and calculate AUC.

```{r}
# format actual outcome values so died = TRUE
outcome <- ifelse(lassa$clinical_outcome.outcome == "died", TRUE, FALSE)
outcome <- factor(outcome, levels = c(TRUE,FALSE))

# format predicted outcomes so p > 0.5 = TRUE
outcome.pred <- ifelse(p > 0.5, TRUE, FALSE)
outcome.pred <- factor(outcome.pred, levels = c(TRUE, FALSE))

# model 1 statistics 
confusionMatrix(outcome.pred, outcome)
```


Calculate AUC

```{r}
roc_pred <- prediction(predictions = p  , labels = outcome)
roc_perf <- performance(roc_pred , "tpr" , "fpr")

auc_ROCR <- performance(roc_pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
```

Plot ROC

```{r}
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))

dir.create("lassa_fever_analysis")
pdf("./lassa_fever_analysis/ROC_model_validation_1_lab-full.pdf")
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))
dev.off()
```

###### Test model #2 (lab-full-HCT)


```{r}
newdata <- drop_na(newdata_all, c(age, ast, bun, k, hct, plt, lym))
intercept <- -7.7053943935 
age <- 0.0320132683
ast <- 0.0014895123
bun <- 0.0290322460 
k <- 0.4141768473 
hct <- 0.0582328918 
plt <- 0.0007336732 
lym <- -0.0006977108 

age_val <- unlist(map(newdata$age, function(x) x * age))
ast_val <- unlist(map(newdata$ast, function(x) x * ast))
bun_val <- unlist(map(newdata$bun, function(x) x * bun))
k_val <- unlist(map(newdata$k, function(x) x * k))
hct_val <- unlist(map(newdata$hct, function(x) x * hct))
plt_val <- unlist(map(newdata$plt, function(x) x * plt))
lym_val <- unlist(map(newdata$lym, function(x) x * lym))

coef <- cbind.data.frame(age_val, ast_val, bun_val, k_val, hct_val, plt_val, lym_val)

coef_sum <- rowSums(coef, na.rm = TRUE)
coef_sum <- coef_sum + intercept
```

```{r}
num <- exp(coef_sum)
denom <- 1 + exp(coef_sum)
p <- num/denom
```

Graph ROC and calculate AUC.

```{r}
# format actual outcome values so died = TRUE
outcome <- ifelse(lassa$clinical_outcome.outcome == "died", TRUE, FALSE)
outcome <- factor(outcome, levels = c(TRUE,FALSE))

# format predicted outcomes so p > 0.5 = TRUE
outcome.pred <- ifelse(p > 0.5, TRUE, FALSE)
outcome.pred <- factor(outcome.pred, levels = c(TRUE, FALSE))

# model 2 statistics 
confusionMatrix(outcome.pred, outcome)
```


Calculate AUC

```{r}
roc_pred <- prediction(predictions = p  , labels = outcome)
roc_perf <- performance(roc_pred , "tpr" , "fpr")

auc_ROCR <- performance(roc_pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
```

Plot ROC

```{r}
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))

dir.create("lassa_fever_analysis")
pdf("./lassa_fever_analysis/ROC_model_validation_2_lab-full-HCT.pdf")
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))
dev.off()
```

###### Test model #3 (lab-full-WBC)

```{r}
newdata <- drop_na(newdata_all, c(age, ast, cre, k, tbil, plt, wbc))
intercept <- -7.632576915
age <- 0.038183853 
ast <- 0.002634452 
cre <- 0.165733651 
k <- 0.804933574 
tbil <- -0.002499645
plt <- 0.001121809 
wbc <- 0.020097665 

age_val <- unlist(map(newdata$age, function(x) x * age))
ast_val <- unlist(map(newdata$ast, function(x) x * ast))
cre_val <- unlist(map(newdata$cre, function(x) x * cre))
k_val <- unlist(map(newdata$k, function(x) x * k))
tbil_val <- unlist(map(newdata$tbil, function(x) x * tbil))
plt_val <- unlist(map(newdata$plt, function(x) x * plt))
wbc_val <- unlist(map(newdata$wbc, function(x) x * wbc))

coef <- cbind.data.frame(age_val, ast_val, cre_val, k_val, tbil_val, plt_val, wbc_val)

coef_sum <- rowSums(coef, na.rm = TRUE)
coef_sum <- coef_sum + intercept
```

```{r}
num <- exp(coef_sum)
denom <- 1 + exp(coef_sum)
p <- num/denom
```

Graph ROC and calculate AUC.

```{r}
# format actual outcome values so died = TRUE
outcome <- ifelse(lassa$clinical_outcome.outcome == "died", TRUE, FALSE)
outcome <- factor(outcome, levels = c(TRUE,FALSE))

# format predicted outcomes so p > 0.5 = TRUE
outcome.pred <- ifelse(p > 0.5, TRUE, FALSE)
outcome.pred <- factor(outcome.pred, levels = c(TRUE, FALSE))

# model 3 statistics 
confusionMatrix(outcome.pred, outcome)
```


Calculate AUC

```{r}
roc_pred <- prediction(predictions = p  , labels = outcome)
roc_perf <- performance(roc_pred , "tpr" , "fpr")

auc_ROCR <- performance(roc_pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
```

Plot ROC

```{r}
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))

dir.create("lassa_fever_analysis")
pdf("./lassa_fever_analysis/ROC_model_validation_3_lab-full-WBC.pdf")
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))
dev.off()
```

###### Test model #4 (lab-min)

```{r}
newdata <- drop_na(newdata_all, c(age, ast, cre, k))
intercept <- -7.65790869
age <- 0.03383892
ast <- 0.00235515 
cre <- 0.14720373 
k <- 0.74449735 

age_val <- unlist(map(newdata$age, function(x) x * age))
ast_val <- unlist(map(newdata$ast, function(x) x * ast))
cre_val <- unlist(map(newdata$cre, function(x) x * cre))
k_val <- unlist(map(newdata$k, function(x) x * k))

coef <- cbind.data.frame(age_val, ast_val, cre_val, k_val)

coef_sum <- rowSums(coef, na.rm = TRUE)
coef_sum <- coef_sum + intercept
```

```{r}
num <- exp(coef_sum)
denom <- 1 + exp(coef_sum)
p <- num/denom
```

Graph ROC and calculate AUC.

```{r}
# format actual outcome values so died = TRUE
outcome <- ifelse(lassa$clinical_outcome.outcome == "died", TRUE, FALSE)
outcome <- factor(outcome, levels = c(TRUE,FALSE))

# format predicted outcomes so p > 0.5 = TRUE
outcome.pred <- ifelse(p > 0.5, TRUE, FALSE)
outcome.pred <- factor(outcome.pred, levels = c(TRUE, FALSE))

# model 4 statistics 
confusionMatrix(outcome.pred, outcome)
```


Calculate AUC

```{r}
roc_pred <- prediction(predictions = p  , labels = outcome)
roc_perf <- performance(roc_pred , "tpr" , "fpr")

auc_ROCR <- performance(roc_pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
```

Plot ROC

```{r}
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))

dir.create("lassa_fever_analysis")
pdf("./lassa_fever_analysis/ROC_model_validation_4_lab-min.pdf")
plot(roc_perf,
     colorize = TRUE)
rect(0, 1.1, 1, 1.7, xpd=TRUE, col="white", border="white")
title(paste0("AUC = ",round(auc_ROCR,2)))
dev.off()
```